---
title: An R Markdown for environmental-footprints
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Context:

As the world's population has expanded and gotten richer, the demand for food, energy and water has seen a rapid increase. Not only has demand for all three increased, but they are also strongly interlinked: food production requires water and energy; traditional energy production demands water resources; agriculture provides a potential energy source. This project focuses on the environmental impacts of food. Ensuring everyone in the world has access to a nutritious diet in a sustainable way is one of the greatest challenges we face.

# Questions:

-   Which types of food have more negative impact on the environment?
-   What types of food production should be encouraged to consume nutritious diet in a sustainable way?
-   Which stage of food production contributes more to the greenhouse gas emission?
-   Compare carbon footprint of plant-based foods?
-   Compare carbon footprint of animal-based foods?
-   Compare carbon footprint of protein rich foods?

```{r}
library(tidyverse)
```

# 1. Loading data from a CSV file

```{r}
data<-read_csv("./input/environment-impact-of-food-production/Food_Production.csv")
head(data)
```

# 2. Cleaning the data

Investigating the structure of data

```{r}
str(data)
```

replacing spaces in column names with '\_'

```{r}
names(data)
names(data)<- gsub("[[:space:]+]", "_", names(data))
names(data)
```

check the data types inferred and convert them if necessary

```{r}
data <- type_convert(data)
#chocolateData$Cocoa_Percent <- sapply(chocolateData$Cocoa_Percent, function(x) gsub("%", "", x))
```

# 3. Exploring the data

```{r}
summary(data)
```

```{r}
summarise_all(data, funs(mean))
```

```{r}
# filtering and exploring some columns and rows in the data

#data %>%
    #group_by() %>%
#    summarise(averageFeed = mean(Land_use_change),
#             FeedSD = sd(Land_use_change))
#ncol(data)
#data%>%select(Animal_Feed) %>% summary()
#data[,c("Animal_Feed","Farm")] %>%summary()
#data.filter(Product="Rice")

#data%>%filter(data['Food product']=='Rice')
#head(data['Processing']) # or data[5]
#data[1,1] #first row first column 
#data[1,] #first row
#data[-1,] #everything except first row
```

# 4. Answering questions

**Which types of food have more negative impact on the environment?**

```{r}
data %>% arrange(desc(Total_emissions)) %>% head()
```

coffee and chocolate are in the top ten products that produce the highest amount of emission , however, they do not provide any calories but their production use so much resources, so measurements such as `Scarcity-weighted_water_use_per_1000kcal_(liters_per_1000_kilocalories)` is very high for such products. Thus, the following question comes up:

**What types of food production should be encouraged to consume nutritious diet in a sustainable way?**

```{r}
data %>% arrange(Total_emissions) %>% head(10)
```

```{r}
water<- "Scarcity-weighted_water_use_per_1000kcal_(liters_per_1000_kilocalories)"
co2<-"Greenhouse_gas_emissions_per_1000kcal_(kgCO₂eq_per_1000kcal)"
land<-"Land_use_per_1000kcal_(m²_per_1000kcal)"
waste<-"Freshwater_withdrawals_per_1000kcal_(liters_per_1000kcal)"
data %>% arrange(desc(water), desc(co2),desc(land), desc(waste)) %>% head(10)
```

```{r}
water<- "Scarcity-weighted_water_use_per_100g_protein_(liters_per_100g_protein)"
co2<-"Greenhouse_gas_emissions_per_100g_protein_(kgCO₂eq_per_100g_protein)"
land<-"Land_use_per_100g_protein_(m²_per_100g_protein)"
waste<-"Freshwater_withdrawals_per_100g_protein_(liters_per_100g_protein)"
data %>% arrange(desc(water), desc(co2), desc(land), desc(waste)) %>% head(10)
```

```{r}
plt<-ggplot(data=data, mapping=aes(y=Food_product, x=Total_emissions))+
geom_col()+
theme_minimal()+
ggtitle("Emission of Food Products")

# save our plot
ggsave("product_emmisions.png", # the name of the file where it will be save
       plot = plt, # what plot to save
       height=6, width=10, units="in")
plt
```

**Which stage of food production contributes more to the greenhouse gas emission?**

```{r}
 
ggplot(data=data)+
geom_point(mapping=aes(x=Farm, y=Total_emissions), color="green") +
geom_point(mapping=aes(x=Land_use_change, y=Total_emissions), color="yellow") +
geom_point(mapping=aes(x=Animal_Feed, y=Total_emissions), color="red") +
geom_point(mapping=aes(x=Transport, y=Total_emissions), color="blue") +
geom_point(mapping=aes(x=Processing, y=Total_emissions), color="black") +
geom_point(mapping=aes(x=Retail, y=Total_emissions), color="orange")
```

```{r}
#install.packages("reshape2")
library(reshape2)
cols<-c("Total_emissions", "Land_use_change","Animal_Feed","Farm","Processing","Transport","Packging","Retail")
corr_mat <- round(cor(data[cols]),2)
corr_mat
 
# reduce the corr matrix
melted_corr_mat <- melt(corr_mat)
 
# plotting the corr heatmap
ggplot(data = melted_corr_mat, aes(x=Var1,
                                   y=Var2,
                                   fill=value)) +
  geom_tile()
```

Farming, Processing and Land use change has the highest emission contribution

Hypothesis : the plant based food produce less emission

```{r}
qqnorm(data$Total_emissions)
```

```{r}
new_data<-data%>% mutate(Plant_based=Animal_Feed<=0)
new_data
```

```{r}
t.test(new_data$Total_emissions ~ new_data$Plant_based)
```

```{r}
ggplot(data = new_data, aes(x = Total_emissions, fill = Plant_based)) +
geom_histogram(binwidth=10)
```
